Our Galactic Midplane Is Probably Wrong, and Not by a Little

### 2018-12-19 paper/ms.tex

\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\z}{z}
\newcommand{\uth}{\textsuperscript{th}} % why is this so long in the first place??

@@

Underlying all of these discoveries is the assumption of a galactocentric
coordinate system \citep{2008gady.book.....B}. In angular
coordinates, the center of the galaxy is taken to be the location of
Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
motions near Sgr~A\textsuperscript{*}, the distance to the center is taken to
be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018A&A...615L..15G}. The vertical height is found
by identifying where the stellar positions and velocities reach a maximum
(effectively the median height of all disk stars), and is usually taken to be
$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.

These measurements of the galactocentric coordinate system rely on the
assumption of an axisymmetric Milky Way. Under this assumption, these
estimates would provide the correct parameters for the galactocentric
coordinate system up to measurement error. However, the Milky Way is not
axisymmetric in detail. Spiral arms, bar(s), and infalling satellite galaxies
such as Sagittarius and the Large Magellanic Cloud all induce
non-axisymmetries.

Since the Milky Way is not truly axisymmetric, it is not clear the best way to
define the Galactocentric coordinate system. In this work, we focus on the midplane and, equivalently, the solar height.

@@

@@

>>>>

\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\z}{z}
\newcommand{\uth}{\textsuperscript{th}}

@@

Underlying all of these discoveries is the assumption of a Galactocentric
coordinate system \citep{2008gady.book.....B}. In angular
coordinates, the center of the galaxy is taken to be the location of
Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
motions near Sgr~A\textsuperscript{*}, the distance to the center is taken to
be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018A&A...615L..15G}. The vertical height is found
by identifying where the stellar positions and velocities reach a maximum
(effectively the median height of all disk stars), and is usually taken to be
$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.

These measurements of the galactocentric coordinate system rely on the
assumption of an axisymmetric Milky Way. Under this assumption, these
estimates would provide the correct parameters for the Galactocentric
coordinate system up to measurement error. However, the Milky Way is not
axisymmetric in detail. Spiral arms, bar(s), and infalling satellite galaxies
such as Sagittarius and the Large Magellanic Cloud all induce
non-axisymmetries.

Since the Milky Way is not truly axisymmetric, it is not clear the best way to
define the Galactocentric coordinate system --- or, perhaps, the best way to
modify the current implicit definition of the Galactocentric coordinate system
in order to account for non-axisymmetries.

@@

The transformation to action space is not trivial, and many of the assumptions
of the transformation are potentially incorrect. It is therefore natural to
consider the reason to transform to action space. One reason is dimensionality
reduction --- in the action space paradigm, a stellar orbit is completely
described by three actions, as opposed to six dimensions of phase space. This
means that the relationship between {\em orbital} properties of individual
stars and stellar properties such as age or metallicity can be investigated.

@@

\section{Discussion} \label{sec:discussion}
{\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
phase space measurements have been made for an unprecedented number of stars.
In the face of such large data, it is worth considering the goal of such
measurements --- a goal which should inform the usage of such data.

In this work we considered one such question: given the current position in
phase space of a single star, what can we infer about that star's past phase
space trajectory? Current methodology involves using a smoothed, analytic form
for the Galactic potential (with perhaps crude implementations of spiral arms
or sattelites), fitting such form to some selection of observations, and
integrating from the currently measured position in phase space. This orbit
integration is also commonly accompanied with a transformation into
action-angle space. Other than the integration, every step of that process
should be viewed with significant scrutiny.

The more general problem of measuring the orbits of stars in cosmological
simulations is essentially unexplored in this work and the literature, and we
delay discussion to future work.
